High-accuracy neutron diffusion calculations based on integral transport theory

### 2019-12-12 docs/EPJP2019/response/response.tex

\newpage

% --------------------------------------------------------------------------------------
\textit{At line 50, please change "w/" with "with". At line 55, please change "w/o" with "without". The use of such contractions (such as "\&" above) is not advisable in a scientific paper. I suggest to make these changes throughout the whole paper.}
\\ \\
\revised{Fixed.}
\\ \\
% --------------------------------------------------------------------------------------
\textit{At line 60, it is advisable to expand on "the corrected diffusion current exhibit(s)non-trivial behaviour near the boundary". Can it be related to BC?}
\\ \\
\revised{Fig. 5 and any reference to it are omitted from the revised manuscript.}
\\ \\
% --------------------------------------------------------------------------------------
\textit{General comment on the Figures: it would be nice to explicitly identify the quantities in the graphs rather than letting the reader struggle to interpret them from the legend.}
\\ \\
\revised{We have added (where possible) explicit text indications on the figure itself to distinguish between the different quantities.}
\\ \\
% --------------------------------------------------------------------------------------
\textit{In Fig. 8: is there any explanation for the sign change in some curves?}
\\ \\
\revised{The following text was added to the manuscript:\\ The initial diffusion solution is smoother than the transport one. Hence, as can be seen in Fig. 4 (in the original and revised ms), the diffusion solution underestimates the peak flux at the center of the slab and overestimates the flux near the boundary. The Ronen iterations "push" the diffusion solution towards the transport one by increasing the flux level at the center of the slab and decreasing the flux level at the boundary. Hence, $\phi^{k-1}$ is smaller in the center and larger at the boundary comparing to $\phi^(k)$.	This effect of the Ronen iterations produces less smooth flux shape with steeper gradients comparing to the diffusion solution. }
\\ \\
% --------------------------------------------------------------------------------------
\textit{Suggestion for further work: it should not be too difficult to study a heterogeneous configuration, since the Ronen method naturally leads to a non-homogeneous system, to see how the Ronen method adapts to interfaces, where strong gradients of the flux might appear.}
\\ \\
\revised{Indeed, the Ronen method is highly suited for non-homogeneous systems. An entire benchmark was added to the results section. The chosen benchmark is a one-dimensional highly heterogeneous core representing a BWR core. As nicely pointed out by the reviewer, the new figures 11-13 in the revised manuscript demonstrate the natural capability of the Ronen method to deal with material interfaces and strong flux gradients.}
\\ \\

### 2019-12-12 docs/EPJP2019/Gross-EPJP-2019.tex
Final version (except minor changes)

Table~\ref{tab:dx_drho} shows the reactivity differences as a function of mesh refinement using standard diffusion (with extrapolated length) and the Ronen method. The reactivity difference is given by $(\Delta\rho = 1/\kref - 1/k_\textrm{D/RM})\times 10^5$ pcm, where $\kd$ and $\krm$ are the multiplication factors of standard diffusion and of the Ronen method, respectively. The reference $\kref$ value from S$_{16}$ is 0.744417 ($I = 400$).

@@

The RM correction terms ($\delta D$) along the slab are shown in Fig.~\ref{fig:slab-RM-corr-factor}. It is possible to observe that the diffusion correction almost vanishes at the slab center and exhibits non-trivial behavior near the boundary. The convergence of the flux (max deviation) and of the criticality eigenvalue are shown in Fig.~\ref{fig:conv}, where $\epsilon_\phi = \operatorname{max} [\phi^{(k)}-\phi^{(k-1)})/\phi^{(k-1)}]$ and $\epsilon_k = (k_\textrm{eff}^{(k)}-k_\textrm{eff}^{(k-1)})/k_\textrm{eff}^{(k-1)}$. While the eigenvalue converges within a few iterations, relatively slow convergence is noticed for the flux with higher differences against the reference S$_{16}$ solution near the boundaries. The spatial flux convergence between two successive Ronen iterations is shown in Fig.~\ref{fig:conv2} for the left half slab, according to $\Delta \phi = (\phi^{k-1}-\phi^k)/\phi^{k-1}$ [\%]. The spatial flux convergence of the Ronen iterations with respect to the reference solution for the left half slab is shown in Fig.~\ref{fig:conv3} instead.

@@

% convergence
\begin{figure}[hbtp]
	\centering
	\includegraphics[width=0.65\linewidth]{RM_Sn_converg_FAST.pdf}
	\includegraphics[width=0.65\linewidth]{RM_Sn_converg_THERMAL.pdf}
	\caption{The spatial flux convergence of the Ronen iterations with respect to the reference S$_{16}$ solution for the left half slab.}
	\label{fig:conv3}
\end{figure}

@@

\begin{table}[htbp]
	\centering
	\caption{Reactivity differences as a function of number of iterations using the Ronen method. 728 nodes.}
	\label{tab:dx_drho2}
	\begin{tabular}{lcccccr}
		%\midrule
		Method & \multicolumn{2}{c}{Core 1} & \multicolumn{2}{c}{Core 2} & \multicolumn{2}{c}{Core 3}  \\ 
		\midrule
		{}				& $\keff$& $\Delta\rho$ [pcm]& $\keff$	&$\Delta\rho$ [pcm]& $\keff$ &$\Delta\rho$ [pcm] \\
		S$_{16}$		&1.25614	&-			 &1.00094   &-				  &0.79878		&-		\\
		Diffusion		&1.25902	&182		 &0.99325	&-773			  &0.77805		&-3335	\\
		RM-10			&1.25677	&40			 &1.00107	&13			 	  &0.79824		&-85	\\
		RM-20			&1.25648	&22			 &1.00091	&-3			  	  &0.79837		&-64	\\
		RM-50			&1.25625	&7			 &1.00076	&-18			  &0.79839		&-61	\\
		RM-100			&1.25618	&3			 &1.00070	&-24			  &0.79838		&-62	\\		
	\end{tabular}
\end{table}

The two-group flux distribution across each core is shown in Fig.~\ref{fig:hetero-flux} for standard diffusion ($D_0$), Ronen method (RM) and the S$_{16}$ reference solution. RM results are shown for 100 iterations and the spatial mesh spans 8 nodes in the water and 22 nodes in each fuel element. On a first glance, it is easy to see that standard diffusion fails to produce the steep flux gradients across material interfaces, especially between fuel and water. This is more pronounced for the fast flux, where even in core 1 it is not described correctly. 
%
\begin{figure}[htbp!]
	\centering
	\includegraphics[width=0.49\linewidth]{Sn_Diff_RM_core_1.pdf}
	\includegraphics[width=0.49\linewidth]{Sn_Diff_RM_core_2.pdf}
	\includegraphics[width=0.49\linewidth]{Sn_Diff_RM_core_3.pdf}
	\caption{Fast and thermal flux for each core using standard diffusion ($D_0$), Ronen method (RM) and S$_{16}$ reference solution. Vertical grid lines are located at assemblies interfaces. RM results are for 100 iterations and the spatial mesh spans 8 nodes in the water and 22 nodes in each fuel element.}
	\label{fig:hetero-flux}
\end{figure}

The fast and thermal flux deviation of standard diffusion ($D_0$) and Ronen method (RM) with respect to S$_{16}$ reference solutions are shown for each core in Fig.~\ref{fig:hetero-flux-dev}. The fast flux is poorly reproduced by standard diffusion, with up to $\sim$10\% deviations at the inner interfaces and $\sim$20\% at the outer boundaries (in agreement with the homogeneous case). Deviations of the thermal flux calculated by diffusion increase from $\sim$4\% to $\sim$10\% as the heterogeneity of the core increases. The deviations of the fast and thermal flux calculated by Ronen method are much smaller, with practically negligible values in the fuel and less than $\sim$2\% in the water.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.75\linewidth]{flx_err_diff_C123.pdf}
	\caption{Fast and thermal flux deviation of standard diffusion ($D_0$) and Ronen method (RM) with respect to S$_{16}$ reference solutions for each core. Vertical grid lines are located at assemblies interfaces. RM results are for 100 iterations and the spatial mesh spans 8 nodes in the water and 22 nodes in each fuel element.}
	\label{fig:hetero-flux-dev}
\end{figure}

@@

More accurate results are obtained for the two-group benchmark problem reported in \cite{Tomatis-2011} and new results are reported for a heterogeneous benchmark. In general, slow convergence is observed for the scalar flux with the larger discrepancies, comparing to the reference, located on the vacuum boundary and on the material interfaces. Although the method is converging to the reference results provided by a discrete ordinate transport code, the improvement of the convergence rate and the use of coarser meshes are crucial for the advancement of the methodology in practical applications. These topics will be addressed as future developments, as well as higher-order anisotropy.

>>>>

Table~\ref{tab:dx_drho} shows the reactivity differences as a function of mesh refinement using standard diffusion (with extrapolated length) and the Ronen method. The reactivity difference is given by $(\Delta\rho = 1/\kref - 1/k_\textrm{D/RM})\times 10^5$ pcm, where $\kd$ and $\krm$ are the multiplication factors of standard diffusion and of the Ronen method, respectively. The reference $\kref$ value from S\tsub{16} is 0.744417 ($I = 400$).

@@

The RM correction terms ($\delta D$) along the slab are shown in Fig.~\ref{fig:slab-RM-corr-factor}. It is possible to observe that the diffusion correction almost vanishes at the slab center and exhibits non-trivial behavior near the boundary. The convergence of the flux (max deviation) and of the criticality eigenvalue are shown in Fig.~\ref{fig:conv}, where $\epsilon_\phi = \operatorname{max} [\phi^{(k)}-\phi^{(k-1)})/\phi^{(k-1)}]$ and $\epsilon_k = (k_\textrm{eff}^{(k)}-k_\textrm{eff}^{(k-1)})/k_\textrm{eff}^{(k-1)}$. While the eigenvalue converges within a few iterations, relatively slow convergence is noticed for the flux with higher differences against the reference S\tsub{16} solution near the boundaries. The spatial flux convergence between two successive Ronen iterations is shown in Fig.~\ref{fig:conv2} for the left half slab, according to $\Delta \phi = (\phi^{k-1}-\phi^k)/\phi^{k-1}$ [\%]. The spatial flux convergence of the Ronen iterations with respect to the reference solution for the left half slab is shown in Fig.~\ref{fig:conv3} instead.

@@

% convergence
\begin{figure}[hbtp]
	\centering
	\includegraphics[width=0.65\linewidth]{RM_Sn_converg_FAST.pdf}
	\includegraphics[width=0.65\linewidth]{RM_Sn_converg_THERMAL.pdf}
	\caption{The spatial flux convergence of the Ronen iterations with respect to the reference S\tsub{16} solution for the left half slab.}
	\label{fig:conv3}
\end{figure}

@@

\begin{table}[htbp]
	\centering
	\caption{Reactivity differences as a function of number of iterations using the Ronen method. 728 nodes.}
	\label{tab:dx_drho2}
	\begin{tabular}{lcccccr}
		%\midrule
		Method & \multicolumn{2}{c}{Core 1} & \multicolumn{2}{c}{Core 2} & \multicolumn{2}{c}{Core 3}  \\ 
		\midrule
		{}				& $\keff$& $\Delta\rho$ [pcm]& $\keff$	&$\Delta\rho$ [pcm]& $\keff$ &$\Delta\rho$ [pcm] \\
		S\tsub{16}		&1.25614	&-			 &1.00094   &-				  &0.79878		&-		\\
		Diffusion		&1.25902	&182		 &0.99325	&-773			  &0.77805		&-3335	\\
		RM-10			&1.25677	&40			 &1.00107	&13			 	  &0.79824		&-85	\\
		RM-20			&1.25648	&22			 &1.00091	&-3			  	  &0.79837		&-64	\\
		RM-50			&1.25625	&7			 &1.00076	&-18			  &0.79839		&-61	\\
		RM-100			&1.25618	&3			 &1.00070	&-24			  &0.79838		&-62	\\		
	\end{tabular}
\end{table}

The two-group flux distribution across each core is shown in Fig.~\ref{fig:hetero-flux} for standard diffusion ($D_0$), Ronen method (RM) and the S\tsub{16} reference solution. RM results are shown for 100 iterations and the spatial mesh spans 8 nodes in the water and 22 nodes in each fuel element. On a first glance, it is easy to see that standard diffusion fails to produce the steep flux gradients across material interfaces, especially between fuel and water. This is more pronounced for the fast flux, where even in core 1 it is not described correctly. 
%
\begin{figure}[htbp!]
	\centering
	\includegraphics[width=0.49\linewidth]{Sn_Diff_RM_core_1.pdf}
	\includegraphics[width=0.49\linewidth]{Sn_Diff_RM_core_2.pdf}
	\includegraphics[width=0.49\linewidth]{Sn_Diff_RM_core_3.pdf}
	\caption{Fast and thermal flux for each core using standard diffusion ($D_0$), Ronen method (RM) and S\tsub{16} reference solution. Vertical grid lines are located at assemblies interfaces. RM results are for 100 iterations and the spatial mesh spans 8 nodes in the water and 22 nodes in each fuel element.}
	\label{fig:hetero-flux}
\end{figure}

The fast and thermal flux deviation of standard diffusion ($D_0$) and Ronen method (RM) with respect to S\tsub{16} reference solutions are shown for each core, in Fig.~\ref{fig:hetero-flux-dev}. The fast flux is poorly reproduced by standard diffusion, with up to $\sim$10\% deviations at the inner interfaces and $\sim$20\% at the outer boundaries (in agreement with the homogeneous case). Deviations of the thermal flux calculated by diffusion increase from $\sim$4\% to $\sim$10\% as the heterogeneity of the core increases. The deviations of the fast and thermal flux calculated by Ronen method are much smaller, with practically negligible values in the fuel and less than $\sim$2\% in the water.

\begin{figure}[htbp]
	\centering
	\includegraphics[width=0.75\linewidth]{flx_err_diff_C123.pdf}
	\caption{Fast and thermal flux deviation of standard diffusion ($D_0$) and Ronen method (RM) with respect to S\tsub{16} reference solutions for each core. Vertical grid lines are located at assemblies interfaces. RM results are for 100 iterations and the spatial mesh spans 8 nodes in the water and 22 nodes in each fuel element.}
	\label{fig:hetero-flux-dev}
\end{figure}

@@

More accurate results are obtained for the two-group benchmark problem reported in \cite{Tomatis-2011} and new results are reported for a heterogeneous benchmark. The performances of the Ronen method, especially for the heterogeneous benchmark (as shown in Figs.~\ref{fig:hetero-flux}--\ref{fig:Dcoef}), strongly indicate that it is inherently suitable to deal with non-homogeneous systems and can naturally adapt to interfaces, where strong gradients of the flux might appear.

In general, slow convergence is observed for the scalar flux with the larger discrepancies, comparing to the reference, located on the vacuum boundary and on the material interfaces. Although the method is converging to the reference results provided by a discrete ordinate transport code, the improvement of the convergence rate and the use of coarser meshes are crucial for the advancement of the methodology in practical applications. These topics will be addressed as future developments, as well as higher-order anisotropy.
